LME-MPM applied to quasi-brittle fracture.

### 2020-06-05 2020_EFM_MPM_Eigensoftening.tex
Cambios 5-5-2020
Casos I y II terminados. Sólo falta completar la simulación del caso III.

The geometry and boundary conditions are sketched in Figure 
\ref{fig:geometry-Semicircular-bending-test}. The experiment consist in two
semicircular specimens with a radius of 47.5 mm and a thickness of
20 mm are supported by two rollers at a span of 47.5 mm and loaded by
another roller on its mid-span. Notches with a length of 16.6 mm are
created with different notch angles of 0$^º$ and 60$^º$ with respect
to the vertical axis to investigate the influences of the
notch angles on the peak load.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}
  \centering
  \includegraphics[width=0.6\textwidth]{./Figure-Semicircular-bending-test}
  \caption{Geometry and boundary condition of the semicircular bending
    test.}
  \label{fig:geometry-Semicircular-bending-test}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
The material properties of Johnstone are listed in Table \ref{tab:Johnstone-properties}.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{table}
  \centering
  \begin{tabular}[]{l l}
    \hline
    Young's modulus (\gls{E})   & 0.4\ $GPa$       \\
    Poisson ratio (\gls{nu})    & 0.25           \\
    Density (\gls{rho})         & 1.54\ $g/cm^3$ \\
    Tensile strength (\gls{ft}) & 0.6\ $MPa$       \\
    Critical opening displacement (\gls{Wc}) & 0.015\ $mm$ \\
    \hline
  \end{tabular}
  \caption[Mechanical properties of Johnstone. ]{Material properties of Johnstone.}
  \label{tab:Johnstone-properties}
\end{table}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

In the tests, particles were uniformly distributed with an almost
constant separation of 1 mm between them, resulting in a total number of 4672
\acrshort{mpm} particles for generating the specimen. Additionally a
regular cartesian background grid of 1917 nodes with 1.5 mm of nodal distance
 was generated.

% The velocities of the top boundary particles are vx =
% 0,vy = − 2 mm/s, whereas the velocities of the bottom boundary
% particles are as vy = 0.

%     \gls{heps}
%     \gls{Ceps}

% As Johnstone rocks are classified as soft rocks with a relatively
% small elastic Young's modulus. Ramp loading is applied to alleviate the dynamic
% effect due to the sudden loading.

>>>>

The geometry and boundary conditions are sketched in Figure 
\ref{fig:geometry-Semicircular-bending-test}. The experiment consist in two
semicircular specimens with a radius of 47.5 mm and a thickness of
20 mm are supported by two rollers at a span of 47.5 mm and loaded by
another roller on its mid-span. Notches with a length of 16.6 mm are
created with different notch angles of 0$^º$ and 60$^º$ with respect
to the vertical axis to investigate the influences of the
notch angles on the peak load \ref{fig:plot-Angle-Forces-test}.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}
  \centering
  \includegraphics[width=0.6\textwidth]{./Figure-Semicircular-bending-test}
  \caption{Geometry and boundary condition of the semicircular bending
    test.}
  \label{fig:geometry-Semicircular-bending-test}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
The material properties of Johnstone are listed in Table \ref{tab:Johnstone-properties}.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{table}
  \centering
  \begin{tabular}[]{l l}
    \hline
    Young's modulus (\gls{E})   & 0.4\ $GPa$       \\
    Poisson ratio (\gls{nu})    & 0.25           \\
    Density (\gls{rho})         & 1.54\ $g/cm^3$ \\
    Tensile strength (\gls{ft}) & 0.6\ $MPa$       \\
    Critical opening displacement (\gls{Wc}) & 0.015\ $mm$ \\
    \hline
  \end{tabular}
  \caption[Mechanical properties of Johnstone. ]{Material properties of Johnstone.}
  \label{tab:Johnstone-properties}
\end{table}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
In the tests, particles were uniformly distributed with an almost
constant separation of 1 mm between them, resulting in a total number of 4672
\acrshort{mpm} particles for generating the specimen. Additionally a
regular cartesian background grid of 1917 nodes with 1.5 mm of nodal distance
was generated. The velocities of the top boundary particles are $v_x$ = 0, $v_y$ = −
2 mm/s, whereas the velocities of the bottom boundary particles are
$v_y$ = 0 mm/s. To reduce dynamic effects due to a sudden loading, it
was preformed within a soft loading ramp. Finally, the model was
calibrated with a regularization parameter \gls{Ceps} = 1.7 and a
bandwidth value \gls{heps} = 15.

Figure \ref{fig:Figure-Angle-Forces-test-damage-pattern} shows the
comparison of mixed-mode fracture envelope between experiments and
\acrshort{mpm} simulations for different notch inclination angles
$\alpha$ = 0$^{\circ}$, 30$^{\circ}$ and 60$^{\circ}$. The final
fracture pattern is a vertical straight line when $\alpha$ =
0$^{\circ}$  (pure mode I). When $\alpha$ > 0$^{\circ}$, the sample is
subjected to mixed-mode. For simulation cases 0$^{\circ}$ and 30$^{\circ}$, the fractures
initiate from the notch tip and then propagate towards upper loading
point. Experiments show that a few fractures tend to initiate behind
notch tip when the notch angle is beyond 50$^{\circ}$, figure
\ref{fig:Figure-Angle-Forces-test-damage-pattern} demonstrate that
eigensoftening combined with \acrshort{mpm} is able to reproduce this
behavior. Therefore, the simulation results predict successfully the
fracture envelope in experiments and demonstrate the capability of
the eigensoftening-\acrshort{mpm} model for capturing the fracture
development under various complex loading conditions. 
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}
\centering
\subfigure[$\alpha$ = 0]{\includegraphics[width=0.49\textwidth]{./Figure-Angle-Forces-test-0-damage-pattern.png}}
\subfigure[$\alpha$ = 30]{\includegraphics[width=0.49\textwidth]{./Figure-Angle-Forces-test-30-damage-pattern.png}}
\subfigure[$\alpha$ = 60]{\includegraphics[width=0.49\textwidth]{./Figure-Angle-Forces-test-60-damage-pattern.png}}
\caption{Crack pattern for different notch inclination angle, and
  comparison with the experimental results  \cite{LIM_1993}.}
\label{fig:Figure-Angle-Forces-test-damage-pattern}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
Additionally, \ref{fig:plot-Angle-Forces-test} compare the peak load
between experiments and simulations for the notch inclination angles
0$^{\circ}$, 30$^{\circ}$ and 60$^{\circ}$. This figure proofs how the
eigensoftening-\acrshort{mpm} simulation can well capture the
increasing trend of the peak load with the increasing notch angles in
the experiment.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}
  \centering
  \includegraphics[width=0.7\textwidth]{./Figure-Angle-Forces-test}
  \caption{Numerical results compared with the experimental results \cite{LIM_1993}.}
  \label{fig:plot-Angle-Forces-test}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2020-06-05 _region_.tex
Cambios 5-5-2020
Casos I y II terminados. Sólo falta completar la simulación del caso III.

The second application to be considered is two semicircular
bending tests on Johnstone, which have been carried out by Lim et
al.95. Wang {\it et al.}\cite{Wang_2020} proposed it with the aim of
examine its predictive capability of the \arcshort{sph} model under a
wide range of mixed-mode loading conditions. Nonetheless, here only
both extreme cases 0$^º$ and 60$^º$ will be reproduced.

The geometry and boundary conditions are sketched in Figure 
\ref{fig:geometry-Semicircular-bending-test} . Similar with 3-point bending test,
the semicircular specimens with a radius of 47.5 mm and a thickness of
20 mm are supported by two rollers at a span of 47.5 mm and loaded by
another roller on its mid-span. Notches with a length of 16.6 mm are
created with different notch angles varying from 0 to 60 with
respect to the vertical axis to investigate the influences of the
notch angles on the peak load (P).

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}
  \centering
  \includegraphics[width=0.6\textwidth]{./Figure-Semicircular-bending-test}
  \caption{Geometry and boundary condition of the semicircular bending
    test.}
  \label{fig:geometry-Semicircular-bending-test}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

The material properties of Johnstone are listed in Table \ref{tab:Johnstone-properties} The friction angle and dilation angle
are respectively taken as 33 and 4  for average quality rock
masses in Hoek and Brown96 for the following examples.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{table}
  \centering
  \begin{tabular}[]{c c c c c c c}
    \hline
      &   $\rho$   & $\text{f}_c$ & $\text{f}_t$ & $G_F$ &   E   & $\text{d}_{max}$ \\
      & (kg/m$^3$) &     (MPa)    &     (MPa)    & (N/m) & (GPa) & (mm) \\
    \hline
Value &    1540    &     102.7    &      5.4     &  141  &   31  &  12 \\
    \hline
  \end{tabular}
  \caption[Mechanical properties of Johnstone. ]{Material properties of Johnstone.}
  \label{tab:Johnstone-properties}
\end{table}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

Figure 9 shows the corresponding SPH configuration of the semicircular
bending test with a notch angle of 30 . In the tests, a particle
discretisation of 0.95 mm is used, resulting in a total number of 3995
SPH particles for generating the specimen. To create accurate notch
angles using reasonable particle numbers, the initial SPH arrangement
in this test is different from the traditional one such as the 3-point
bending test in Section 5.1. The initial location of each SPH particle in semicircular bending tests is related to the notch angle as shown in Figure 9B. Similar to the 3-point bending
tests, three layers of boundary particles are generated for simulating
the point loading in the experiments. The velocities of the top boundary particles are vx = 0,vy = − 2 mm/s, whereas the velocities of the bottom boundary particles are
as vy = 0.

As Johnstone rocks are classified as soft rocks with a relatively
small elastic Young's modulus, both artificial viscosity and
artificial stress are used in this test to stabilise the numerical
solutions. The ramp loading is also applied to alleviate the dynamic
effect due to the sudden loading.

The stresses of boundary particles are updated in the same way as that
in the 3-point bending test. The stresses of real particles near the
loading particles are recorded during the simulation process, from
which the peak load and fracture toughness can be obtained for the
comparison with the experimental data.

Figure 10 illustrates the progressive development of fracture and the
horizontal velocity profile at the failure stage for the semicircular specimen with the rotational angle of 30 . As seen in Figure 10A, some particles near the tip of the
notch are first damaged, showing that the fracture is initiated at the
location of the notch tip. Then, the fracture propagates towards the
loading point (Figure 10B–D), and eventually, a vertical curvilinear
fracture pattern is formed in the middle of the sample (Figure
10E). The horizontal velocity profile corresponding to Figure 10E is
also plotted in Figure 10F. It can be seen that the particles on the
left side of the fracture move to the left whereas the particles on
the other side move to the right direction; thus, the sample is split
into two halves at the end. In this approach, SPH particle itself could not be cracked visually; however, each damaged SPH particle possesses a FPZ, and the material behaviour
within the FPZ will contribute to the decrease in the overall stress
state of the SPH particle, thereby resulting in the separation between
two damaged SPH particles. The same test without using the artificial
stress is also plotted in Figure 11. The result remains stable when
the fracture is initiated (Figure 11A) but becomes unstable during the
post-failure process (Figure 11B,C). Unrealistic fracture and
particles clustering are observed in Figure 11B,C, which indicates the
tensile instability problem is occurred in SPH. By comparing those two
results, it can be concluded that the tensile instability problem can
be completely removed and good numerical stability can be achieved by
just adopting this artificial stress term. Figure 12 depicts the
comparison of mixed-mode fracture envelope between experiments and SPH
simulations for different notch inclination angles α = 0 , 15  ,
30  , 45 , 50 and 60 . It shows that the simulation fracture
patterns marked as red lines match well with the experimental results
shown as black lines. The final fracture pattern is a vertical
straight line when α = 0 (pure mode I). When α > 0 , the sample is
subjected to mixed-mode or pure mode II loading, resulting in a curvilinear fracture paths. For all simulation cases, the fractures initiate from the notch tip and then
propagate towards upper loading point, while the experiments show that
a few fractures tend to initiate behind the notch tip when the notch angle is beyond α =50 . Nevertheless, the simulation results still could successfully predict
the overall fracture envelope in experiments and demonstrate the
capability of the traditional SPH model for capturing the fracture development under various complex loading conditions. In the meanwhile, the comparison of peak load
between experiments and simulations for the notch inclination angles
varying from 0 to 60 with an internal increment of 5 is plotted
in Figure 13A. It is noted here that a notch angle of 54 (instead of
55) is used in the current test as the sample is subjected to pure
mode II loading condition when α = 54 .95 The SPH simulation can
well capture the increasing trend of the peak load with the increasing
notch angles in the experiment. Additionally, Figure 13B compares the
predictions of fracture toughness in SPH simulation against the
experimental results and Lim's theory.95 The simulation results perfectly fall within the experimental data ranges even though the fracture toughness data in the
experiments are quite scattered. Nevertheless, the numerical results
agree closely with the empirical fracture toughness envelope, which
was proposed by Lim et al.95 and represented as a black line Figure
13B. It further confirms that our model could accurately predict the
mixed-mode fracture behaviour of rocks in semicircular bending test.

@@

\message{ !name(2020_EFM_MPM_Eigensoftening.tex) !offset(-1283) }

>>>>

The second application to be considered is two semicircular
bending tests on Johnstone, which have been carried out by Lim {\it et
  al.} \cite{LIM_1993} and numerically reproduced by Wang {\it et
  al.}\cite{Wang_2020} under \acrshort{sph} framework within an
implicit algorithm. The aim of the test is to examine the predictive
capability of the eigensoftening algorithm under different kind of
mixed-mode loading conditions. This capability was previously within a
\acrshort{otm} implementation \cite{Navas_2018_ES}.

The geometry and boundary conditions are sketched in Figure 
\ref{fig:geometry-Semicircular-bending-test}. The experiment consist in two
semicircular specimens with a radius of 47.5 mm and a thickness of
20 mm are supported by two rollers at a span of 47.5 mm and loaded by
another roller on its mid-span. Notches with a length of 16.6 mm are
created with different notch angles of 0$^º$ and 60$^º$ with respect
to the vertical axis to investigate the influences of the
notch angles on the peak load \ref{fig:plot-Angle-Forces-test}.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}
  \centering
  \includegraphics[width=0.6\textwidth]{./Figure-Semicircular-bending-test}
  \caption{Geometry and boundary condition of the semicircular bending
    test.}
  \label{fig:geometry-Semicircular-bending-test}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
The material properties of Johnstone are listed in Table \ref{tab:Johnstone-properties}.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{table}
  \centering
  \begin{tabular}[]{l l}
    \hline
    Young's modulus (\gls{E})   & 0.4\ $GPa$       \\
    Poisson ratio (\gls{nu})    & 0.25           \\
    Density (\gls{rho})         & 1.54\ $g/cm^3$ \\
    Tensile strength (\gls{ft}) & 0.6\ $MPa$       \\
    Critical opening displacement (\gls{Wc}) & 0.015\ $mm$ \\
    \hline
  \end{tabular}
  \caption[Mechanical properties of Johnstone. ]{Material properties of Johnstone.}
  \label{tab:Johnstone-properties}
\end{table}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
In the tests, particles were uniformly distributed with an almost
constant separation of 1 mm between them, resulting in a total number of 4672
\acrshort{mpm} particles for generating the specimen. Additionally a
regular cartesian background grid of 1917 nodes with 1.5 mm of nodal distance
was generated. The velocities of the top boundary particles are $v_x$ = 0, $v_y$ = −
2 mm/s, whereas the velocities of the bottom boundary particles are
$v_y$ = 0 mm/s. To reduce dynamic effects due to a sudden loading, it
was preformed within a soft loading ramp. Finally, the model was
calibrated with a regularization parameter \gls{Ceps} = 1.7 and a
bandwidth value \gls{heps} = 15.

Figure \ref{fig:Figure-Angle-Forces-test-damage-pattern} shows the
comparison of mixed-mode fracture envelope between experiments and
\acrshort{mpm} simulations for different notch inclination angles
$\alpha$ = 0$^{\circ}$, 30$^{\circ}$ and 60$^{\circ}$. The final
fracture pattern is a vertical straight line when $\alpha$ =
0$^{\circ}$  (pure mode I). When $\alpha$ > 0$^{\circ}$, the sample is
subjected to mixed-mode. For simulation cases 0$^{\circ}$ and 30$^{\circ}$, the fractures
initiate from the notch tip and then propagate towards upper loading
point. Experiments show that a few fractures tend to initiate behind
notch tip when the notch angle is beyond 50$^{\circ}$, figure
\ref{fig:Figure-Angle-Forces-test-damage-pattern} demonstrate that
eigensoftening combined with \acrshort{mpm} is able to reproduce this
behavior. Therefore, the simulation results predict successfully the
fracture envelope in experiments and demonstrate the capability of
the eigensoftening-\acrshort{mpm} model for capturing the fracture
development under various complex loading conditions. 
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}
\centering
\subfigure[$\alpha$ = 0]{\includegraphics[width=0.49\textwidth]{./Figure-Angle-Forces-test-0-damage-pattern.png}}
\subfigure[$\alpha$ = 30]{\includegraphics[width=0.49\textwidth]{./Figure-Angle-Forces-test-30-damage-pattern.png}}
\subfigure[$\alpha$ = 60]{\includegraphics[width=0.49\textwidth]{./Figure-Angle-Forces-test-60-damage-pattern.png}}
\caption{Crack pattern for different notch inclination angle, and
  comparison with the experimental results  \cite{LIM_1993}.}
\label{fig:Figure-Angle-Forces-test-damage-pattern}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
Additionally, \ref{fig:plot-Angle-Forces-test} compare the peak load
between experiments and simulations for the notch inclination angles
0$^{\circ}$, 30$^{\circ}$ and 60$^{\circ}$. This figure proofs how the
eigensoftening-\acrshort{mpm} simulation can well capture the
increasing trend of the peak load with the increasing notch angles in
the experiment.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}
  \centering
  \includegraphics[width=0.7\textwidth]{./Figure-Angle-Forces-test}
  \caption{Numerical results compared with the experimental results \cite{LIM_1993}.}
  \label{fig:plot-Angle-Forces-test}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

\message{ !name(2020_EFM_MPM_Eigensoftening.tex) !offset(-1280) }
